---
output: pdf_document
---
<!--Code for importing CSV and JSON files into memory.
      Written by Aaron Niskin on 25 AUG 2016  -->

---
title: "Assignment 2"
author: "Aaron Niskin"
date: "August 25, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
There is a great deal of data available in the form of Extensible Markup Language (xml) files. There is a
good reference for xml files:
https://www.sitepoint.com/really-good-introduction-xml/
Navigate to http://www.cs.washington.edu/research/xmldatasets/www/repository.html#auctions
Prepare an R Markdown document which documents the use of R tools/code to address the following

1. Download the 277KB xml file associated to Reed courses.
```{r, eval = TRUE, include=TRUE}
library(XML)
reed_df <- xmlToDataFrame("http://www.cs.washington.edu/research/xmldatasets/data/courses/reed.xml")
```

2. How many distinct subjects are listed in the document?

```{r, eval=TRUE, include=TRUE}
answer_2 <- length(unique(reed_df[,'subj']))
answer_2
```

3. How many courses have a NULL instructor listing?

```{r, eval=TRUE, include=TRUE}
answer_3 <- length(reed_df[reed_df$instructor == '',][,'instructor'])
answer_3
```

4. How many distinct instructors are listed?

```{r, eval=TRUE, include=TRUE}
answer_4 <- length(unique(reed_df[,'instructor'][reed_df$instructor != '']))
answer_4
```
There are `r answer_4` instructors.
