---
title: "Predicting Violent Crime with Census Data"
author: "Aaron Niskin"
date: "November 10, 2016"
output:
  pdf_document:
    highlight: tango
    toc: yes
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy = TRUE, warning=FALSE, message=FALSE, echo=FALSE, results='hide')
```

# Abstract
Does poverty have a notable effect on arrest rates?

[see @censusB]

```{r, echo=FALSE}
setwd("~/Documents/courses/newCollege/current/eda/assignments/a10/")
library(xlsx)
library(tidyverse)
library(xlsx)
```

# Introduction

# Methods
About the data: [see @pop06 and, @pop12]

# Results

# Discussion

Data was taken from the Florida Department of Law Enforcement, and from the American Community Survey (ACS), data set ID: B20004 organized by the Census Bureau. The data contains: median earnings over a twelve month period by sex and education for those 25 years or older. The table IDs for those Census Bureau tables are ACS_**_1YR_B20004 where the "**" is replaced by the two digit year abbreviation (i.e. ACS_06_1YR_B20004 - ACS_12_1YR_B20004).

Beuraux of labor statistics

```{r, eval = FALSE}
filenames <- list.files("ignore/FDLE", pattern="*.xls", full.names=TRUE)
ldf <- lapply(filenames, function(x) read.xlsx(x, sheetName ="Part II (1)", startRow=3))

#Now to confirm that the structure is regular (with respect to rows).
Alachua_label <- lapply(ldf, function(x) which((x[,1] == "Alachua County") | (x[,1] =="Alachua") ))
Washington_label <- lapply(ldf, function(x) which((x[,1] == "Washington County") | (x[,1] == "Washington") ))
unlist(Washington_label) - unlist(Alachua_label)

# To check if the files all have the same column order:
lapply(ldf, function(x) {return(names(x))})

# Note: I noticed that in the last file, the data starts on column 4 instead of column 3, so I deleted column 3 (and column 5 which was empty).

# Because the different files have different column names
CNames <- names(as.data.frame(ldf[1]))
L<- length(as.data.frame(ldf[1])[1,])
year = 2006
FL_Crime_0612 <- as.data.frame(ldf[1])[Alachua_label[[1]]: Washington_label[[1]],]
FL_Crime_0612$Year = year
for (i in 2:length(ldf)) {
  year = year + 1
  tmp <- as.data.frame(ldf[i])[Alachua_label[[i]]:Washington_label[[i]],1:L]
  names(tmp) <- CNames
  tmp$Year = year
  FL_Crime_0612 <- rbind(FL_Crime_0612, tmp)
  rm(tmp)
}
rm(list = c("i", "year", "L", "CNames", "filenames", "ldf", "Alachua_label", "Washington_label"))

names(FL_Crime_0612)[1] <- "County"
FL_Crime_0612$County <- as.factor(gsub("Miami Dade", "Miami-Dade", FL_Crime_0612$County))
str(FL_Crime_0612)

FL_Crime_0612[,1] <- as.factor(gsub("\\ County", "", FL_Crime_0612[,1]))
unique(FL_Crime_0612$County)
FL_Crime_0612[,1] <- as.factor(gsub("Desoto", "DeSoto", FL_Crime_0612[,1]))
unique(FL_Crime_0612$County)
write.csv(FL_Crime_0612, "ignore/FDLE/FL_Crime_0612.csv",row.names=FALSE)
```

```{r}
FL_Crime_0612 <- read.csv("ignore/FDLE/FL_Crime_0612.csv")
```

```{r, eval=FALSE}
#### To get population by county by year
#### First remove leading periods and any extraneous rows
tmpDf <- read.csv("ignore/population/pop_00-09_edit.csv", stringsAsFactors = FALSE)
names(tmpDf)[1] <- "County"
names(tmpDf) <- gsub("July.1..", "pop_", names(tmpDf))
popYear <- tmpDf[,c("County", "pop_2006")]
popYear <- cbind(popYear, 2006)
names(popYear)[3] <- "Year"
names(popYear)[2] <- "Population"
popYear$Population <- as.numeric(gsub(",", "", popYear$Population))
for(i in 7:9){
  tmp <- tmpDf[,c("County", paste("pop_200", i, sep = ""))]
  tmp <- cbind(tmp, 2000+i)
  names(tmp)[3] <- "Year"
  names(tmp)[2] <- "Population"
  tmp$Population <- as.numeric(gsub(",", "", tmp$Population))
  popYear <- rbind(popYear, tmp)
  rm(tmp)
}
rm(list= c("tmpDf", "i"))
tmpDf <- read.csv("ignore/population/PEP_2012_PEPANNRES_with_ann.csv", stringsAsFactors = FALSE, skip = 1)
tmpDf <- tmpDf[,c(3,6:8)]
names(tmpDf) <- c("County", "pop_2010", "pop_2011", "pop_2012")
for(i in 10:12){
  tmp <- tmpDf[,c("County", paste("pop_20", i, sep = ""))]
  tmp <- cbind(tmp, 2000+i)
  names(tmp)[3] <- "Year"
  names(tmp)[2] <- "Population"
  tmp$Population <- as.numeric(gsub(",", "", tmp$Population))
  tmp$County <- as.factor(gsub("\\ County,\\ Florida", "", tmp$County))
  popYear <- rbind(popYear, tmp)
  rm(tmp)
}
rm(list= c("tmpDf", "i"))
write.csv(popYear, "ignore/population/popYear.csv", row.names = FALSE)
```

```{r}
popYear <- read.csv("ignore/population/popYear.csv")
```


```{r, eval=FALSE}
### For government statistics
govDf <- read.csv("ignore/gov/COG_2012_ORG13.ST05P_with_ann.csv", skip = 1, na.strings = c("-"))
names(govDf)[which(names(govDf) == "Geographic.area..4..1")] <- "County"
govDf$County <- as.factor(gsub("\\ County", "", govDf$County))
govDf <- govDf[, c(7:13, 15:17)]
write.csv(govDf, "ignore/gov/govDf.csv", row.names=FALSE)
```

```{r}
govDf <- read.csv("ignore/gov/govDf.csv")
```

```{r, eval=FALSE}
#### For voting history
tmpDf <- read.xlsx2("ignore/gen2012_countyprecinctparty.xls", sheetName = "Sheet1")
tmpDf$COUNT <- as.integer(as.character(tmpDf$COUNT...))
tmpVoterDf <- select(tmpDf, -PRECINCT_NUM) %>% select(-COUNT...) 
tmpVoterDf <- tmpVoterDf %>% group_by(COUNTY_ID, PARTY_GROUP) %>% summarize(COUNT = sum(COUNT)) %>% mutate(PARTY_GROUP = as.character(PARTY_GROUP)) %>% ungroup() %>% group_by(COUNTY_ID) %>% spread(PARTY_GROUP, COUNT) 
tmpVoterDf <- rename(tmpVoterDf, Abbv = COUNTY_ID)

countyAbbs <- "ALA-Alachua-North	HAM-Hamilton-North	OKA-Okaloosa-Panhandle
BAK-Baker-North	HAR-Hardee-Central	OKE-Okeechobee-Central
BAY-Bay-Panhandle	HEN-Hendry-South	ORA-Orange-Central
BRA-Bradford-North	HER-Hernando-Central	PAL-Palm Beach-South
BRE-Brevard-Central	HIG-Highlands-Central	PAS-Pasco-Central
BRO-Broward-South	HIL-Hillsborough-Central	PIN-Pinellas-Central
CAL-Calhoun-Panhandle	HOL-Holmes-Panhandle	POL-Polk-Central
CHA-Charlotte-Central	IND-Indian River-Central	PUT-Putnam-North
CIT-Citrus-Central	JAC-Jackson-Panhandle	SAR-Sarasota-Central
CLA-Clay-North	JEF-Jefferson-Panhandle	SEM-Seminole-Central
COLL-Collier-South	LAF-Lafayette-North	SJO-St. Johns-North
COLU-Columbia-North	LAK-Lake-Central	SLU-St. Lucie-Central
DAD-Dade-South	LEE-Lee-South	SRO-Santa Rosa -Panhandle
DES-DeSoto-Central	LEO-Leon-Panhandle	SUM-Sumter-Central
DIX-Dixie-North	LEV-Levy-North	SUW-Suwanee-North
DUV-Duval-North	LIB-Liberty-Panhandle	TAY-Taylor-North
ESC-Escambia-Panhandle	MAD-Madison-North	UNI-Union-North
FLA-Flagler-North	MAN-Manatee-Central	VOL-Volusia-Central
FRA-Franklin-Panhandle	MARI-Marion-North	WAK-Wakulla-Panhandle
GAD-Gadsden-Panhandle	MART-Martin-Central	WAL-Walton-Panhandle
GIL-Gilchrist-North	MON-Monroe-South	WAS-Washington-Panhandle
GLA-Glades-Central	NAS-Nassau-North	 
GUL-Gulf-Panhandle	OSC-Osceola-Central"

countyAbbs <- gsub("\\n", ",", gsub("\\t", ",", countyAbbs))
tmp <- strsplit(countyAbbs, ",")[1]
tmp <- as.data.frame(lapply(tmp, function(st) gsub("(.*)-(.*)-.*", "\\1-\\2", st)))
names(tmp) <- "Abbv"
countyAbbvs <- filter(tmp, unlist(1:length(tmp$Abbv) %in% grep("[A-Z]", tmp$Abbv))) %>% separate(Abbv, c("Abbv", "County"), sep="-")
countyAbbvs$County <- as.factor(gsub("Dade", "Miami-Dade", countyAbbvs$County))
countyAbbvs$Abbv <- countyAbbvs$Abbv %>% gsub("COLL", "CLL", .) %>% gsub("COLU", "CLM", .) %>% gsub("SJO", "STJ", .) %>% gsub("SLU", "STL", .) %>% gsub("SRO", "SAN", .) %>% gsub("MARI", "MRN", .) %>% gsub("MART", "MRT", .)

voterDf <- tmpVoterDf %>% full_join(countyAbbvs)
voterDf <- voterDf[,c(5,2,4,3)]
write.csv(countyAbbvs, "ignore/countyAbbvs.csv", row.names=FALSE)
rm(list=c("tmpDf", "countyAbbs", "tmp", "tmpVoterDf", "countyAbbvs"))
write.csv(voterDf, "ignore/voteByCounty.csv", quote = FALSE, row.names = FALSE)
```

```{r}
voterDf = read.csv("ignore/voteByCounty.csv")
```

[see @voter]

http://dos.myflorida.com/elections/data-statistics/voter-registration-statistics/bookclosing/bookclosing-reports-regular/

http://entnemdept.ifas.ufl.edu/teneb/florida_checklist_map.htm
for the abbreviations

```{r, eval=FALSE}
mergedDf <- FL_Crime_0612 %>% inner_join(popYear)
get_ratio <- function(col1, col2){
  tmp <- sapply(1:length(col1), function(i) col1[[i]] * 1.0 / col2[[i]])
  return(tmp)
}
tmpDf <- mergedDf %>% select(2:11) %>% mutate_each(funs(function(col) { return(get_ratio(col, mergedDf$Population)) }))
tmpDf <- mergedDf %>% group_by(County) %>% select(2:11) %>% summarise_each(funs(sum)) %>% mutate_all(funs( function(x) {return(x / mergedDf$Population)}))
str(tmpDf)
govMerged <- tmpDf %>% inner_join(govDf)
str(govMerged)
```



```{r, eval=FALSE}
### The following are for the county demographics data
#### Only some counties are included
filenames <- list.files("ignore/ACS_B20004/", pattern="*ann.csv", full.names=TRUE)

ldf <- lapply(filenames, function(x) read.csv(x, header = TRUE, skip = 1, na.strings = c("-", "***", "*****", "N", "(X)"), stringsAsFactors = FALSE))
startYear = 2006
colList = c(3,4,6,8,10,12,14)
edByC <- ldf[[1]][,colList]
edByC$year <- startYear
for (i in 2:length(ldf)) {
  tmp <- ldf[[i]][,colList]
  # This line is because file number 7 changed the naming scheme
  names(tmp) <- gsub("Total\\.\\.\\.\\.", "", names(tmp))
  tmp$year = startYear + i - 1
  edByC <- rbind(edByC, tmp)
  rm(tmp)
}
rm(list= c("i", "startYear", "colList", 'filenames', "ldf"))
edByC$County <- as.factor(gsub(" County, Florida", "", edByC$Geography))
edByC <- edByC[,which(names(edByC) != "Geography")]
str(edByC)
unique(edByC$County)
fullDf <- inner_join(FL_Crime_0612, edByC, by= c("year", "County"))
```

```{r, eval=FALSE}
ggplot(data=fullDf, aes(x=(Estimate..Bachelor.s.degree / Estimate..Total.), y=(Kidnap..Abduction + Simple.Assault))) + geom_point()

ggplot(data=fullDf, aes(x=(Estimate..Graduate.or.professional.degree / Estimate..Total.), y=Kidnap..Abduction + Simple.Assault)) + geom_point() + facet_grid(year ~ .)

ggplot(data=fullDf, aes(x=Estimate..Graduate.or.professional.degree, y=Kidnap..Abduction + Simple.Assault)) + geom_point() + facet_grid(year ~ .)

ggplot(data=fullDf, aes(x=Estimate..Graduate.or.professional.degree, y=Kidnap..Abduction + Simple.Assault)) + geom_point()

ggplot(data=fullDf, aes(x=(Estimate..Graduate.or.professional.degree / Estimate..Bachelor.s.degree), y=(Kidnap..Abduction + Simple.Assault))) + geom_point()

ggplot(data=fullDf, aes(x=Estimate..Total., y=Kidnap..Abduction + Simple.Assault)) + geom_point()

ggplot(data=fullDf, aes(x = Estimate..Total., y=Estimate..Graduate.or.professional.degree)) + geom_point()

ggplot(data=fullDf, aes(x=Kidnap..Abduction, y=Simple.Assault)) + geom_point()
fullDf[fullDf$Kidnap..Abduction > 100, "County"]
fullDf[fullDf$Simple.Assault > 8000, "County"]

ggplot(data=fullDf, aes(x=Kidnap..Abduction, y=Fraud)) + geom_point()
fullDf[fullDf$Fraud > 1200, "County"]

ggplot(data=fullDf, aes(x=Simple.Assault, y=Fraud)) + geom_point()
fullDf[fullDf$Fraud > 1200, "County"]

ggplot(data=fullDf, aes(x=Simple.Assault, y=Embezzlement)) + geom_point()
fullDf[fullDf$Fraud > 1200, "County"]

ggplot(data=fullDf[fullDf$Kidnap..Abduction < 100,], aes(x=Kidnap..Abduction, y=Arson)) + geom_point()

ggplot(data=fullDf, aes(x=Extortion..Blackmail, y=Embezzlement)) + geom_point()

ggplot(data=fullDf, aes(x=Estimate..Total., y=Drug.Arrest)) + geom_point()
fullDf[fullDf$Drug.Arrest > 10000, "County"]
fullDf[fullDf$County == "Miami-Dade", "County"]

```


# Appendix
Things

## Code

## References
