---
title: 'EDA: Crime In Florida'
author: "P. McDonald"
date: "November 10, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Last time

*   Advice on presentations
*   EDA
    +  Feedback loops
    +  Working with date/time
    +  Finding the correct question

Tukey:

"An approximate answer to the right problem is worth a good deal more than an exact answer to an approximate problem."

http://www.azquotes.com/author/14847-John_Tukey

##Crime in Florida

Crime statistics are maintained by the FBI and are open:

http://www.ucrdatatool.gov/Search/Crime/Local/LocalCrime.cfm

##Crime in Florida

Crime statistics are maintained by the FBI and are open:

http://www.ucrdatatool.gov/Search/Crime/Local/LocalCrime.cfm

The associated tools allow for a variety of parameter selections.

I downloaded eight years of data for Florida: eight tables formatted as csv.

##Crime in Florida

Crime statistics are also maintained by the State of Florida and are open:

http://www.fdle.state.fl.us/cms/FSAC/Data-Statistics/UCR-Arrest-Data.aspx

##Crime in Florida

Crime statistics are also maintained by the State of Florida and are open:

http://www.fdle.state.fl.us/cms/FSAC/Data-Statistics/UCR-Arrest-Data.aspx

I downloaded files containing yearly reports of arrests for a variety of crimes in each county in Florida.

##Crime in Florida

Crime statistics are also maintained by the State of Florida and are open:

http://www.fdle.state.fl.us/cms/FSAC/Data-Statistics/UCR-Arrest-Data.aspx

I downloaded files containing yearly reports of arrests for a variety of crimes in each county in Florida.

I downloaded yearly reports from 2006-2012: seven tables formatted as xslx.

##Crime in Florida

Crime statistics are also maintained by the State of Florida and are open:

http://www.fdle.state.fl.us/cms/FSAC/Data-Statistics/UCR-Arrest-Data.aspx

I downloaded files containing yearly reports of arrests for a variety of crimes in each county in Florida.

I downloaded yearly reports from 2006-2012: seven tables formatted as xslx.

There is some munging to be done.

##Reading files in a directory

It's possible to automate tasks using lists.  We can read the files as a batch:

```{r,eval=FALSE}
filenames <- list.files("~/Classes/16-17/Fall/EDA2016/Data/FDLE", 
                        pattern="*.xls", full.names=TRUE)
library(tidyverse)
library(xlsx)

ldf <- lapply(filenames, function(x) 
  read.xlsx(x, sheetName ="PT2(1)", startRow=3))

CNames <- names(as.data.frame(ldf[1]))
Alachua_label <- lapply(ldf, function(x) 
  which((x[,1] == "Alachua County") | (x[,1] =="Alachua") ))
Washington_label <- lapply(ldf, function(x) 
  which((x[,1] == "Washington County") | (x[,1] == "Washington") ))
```
```{r, echo=FALSE,warning=FALSE, message=FALSE}
filenames <- list.files("~/Classes/16-17/Fall/EDA2016/Data/FDLE", pattern="*.xls", full.names=TRUE)
library(xlsx)
library(tidyverse)
ldf <- lapply(filenames, function(x) read.xlsx(x, sheetName ="PT2(1)", startRow=3))
CNames <- names(as.data.frame(ldf[1]))
Alachua_label <- lapply(ldf, function(x) which((x[,1] == "Alachua County") | (x[,1] =="Alachua") ))
Washington_label <- lapply(ldf, function(x) which((x[,1] == "Washington County") | (x[,1] == "Washington") ))
```
##Checking completeness

It would be good to know that there are a constant number of counties:

```{r}
unlist(Washington_label) - unlist(Alachua_label)
```

##Checking completeness

It would be good to know that there are a constant number of counties:

```{r}
unlist(Washington_label) - unlist(Alachua_label)
```

A good sign:  all the files have the same number of rows: `r unlist(Washington_label)[1] - unlist(Alachua_label)[1]`.

Another good sign:  Florida has 67 counties.

##Naming tricks

Our goal is to concatenate a list of files into a data frame.

##Naming tricks

Our goal is to concatenate a list of files into a data frame.

A simple strategy is to make sure the files have the same number of columns and use rbind.  

##Naming tricks

Our goal is to concatenate a list of files into a data frame.

A simple strategy is to make sure the files have the same number of columns and use rbind.

It turns out that there are other obstructions.  One serious obstruction is column names.  

##Naming tricks

Our goal is to concatenate a list of files into a data frame.

A simple strategy is to make sure the files have the same number of columns and use rbind.

It turns out that there are other obstructions.  One serious obstruction is column names.  

Here is a way to set all column names to NULL
```{r}
NULL_Name <- function(x) {
  names(x) <- NULL
  return(x)
}
ldf <- lapply(ldf, NULL_Name)
```

##Generating a frame:
Now build a data frame  by truncatating each frame and using rbind:

```{r,eval=FALSE}
L<- length(as.data.frame(ldf[1])[1,])
FL_Crime_0612 <- as.data.frame(ldf[1])[Alachua_label[[1]]: 
                                         Washington_label[[1]],]
for (i in 2:length(ldf)) {
  FL_Crime_0612 <- rbind(FL_Crime_0612, 
                         as.data.frame(ldf[i])[Alachua_label[[i]]: 
                                         Washington_label[[i]],1:L])
}

FL_Crime_0612 <- as.data.frame(FL_Crime_0612)
names(FL_Crime_0612) <-CNames
str(FL_Crime_0612)
```
```{r,echo=FALSE}
L<- length(as.data.frame(ldf[1])[1,])
FL_Crime_0612 <- as.data.frame(ldf[1])[Alachua_label[[1]]:Washington_label[[1]],]
for (i in 2:length(ldf)) {
  FL_Crime_0612 <- rbind(FL_Crime_0612, as.data.frame(ldf[i])[Alachua_label[[i]]:Washington_label[[i]],1:L])
}
FL_Crime_0612 <- as.data.frame(FL_Crime_0612)
names(FL_Crime_0612) <-CNames
str(FL_Crime_0612)
```
##Cleaning things up

Let's get rid of the "County" label distinction, fix data types and attach years:

```{r,eval=FALSE}
FL_Crime_0612[,1] <- as.factor(gsub("\\ County", "", FL_Crime_0612[,1]))
FL_Crime_0612[,1] <- as.factor(gsub("Desoto", "DeSoto", FL_Crime_0612[,1]))

FL_Crime_0612[,2:length(FL_Crime_0612[1,])] <-  
  sapply(FL_Crime_0612[,2:length(FL_Crime_0612[1,])], as.integer)

start <- rep(2006,67)
FL_Crime_0612$Year <- c(start,start+1,start+2,start+3,start+4,
                        start+5,start+6)
colnames(FL_Crime_0612)[1] <- "County"
str(FL_Crime_0612)
```
```{r,echo=FALSE}
FL_Crime_0612[,1] <- as.factor(gsub("\\ County", "", FL_Crime_0612[,1]))

FL_Crime_0612[,2:length(FL_Crime_0612[1,])] <-  sapply(FL_Crime_0612[,2:length(FL_Crime_0612[1,])], as.integer)

start <- rep(2006,67)
FL_Crime_0612$Year <- c(start,start+1,start+2,start+3,start+4,start+5,start+6)
colnames(FL_Crime_0612)[1] <- "County"
str(FL_Crime_0612)
```

##Write the file:

Now that we have a clean data frame, write it to disk:

```{r,eval=FALSE}
write.csv(FL_Crime_0612, "~/Classes/16-17/Fall/EDA2016/Data/FDLE/
          FL_Crime_0612.csv")
```
```{r,echo=FALSE}
write.csv(FL_Crime_0612, "~/Classes/16-17/Fall/EDA2016/Data/FDLE/FL_Crime_0612.csv",row.names=FALSE)
```

##Other parameters

We can study the correlation of crime with other factors.

##Other parameters

We can study the correlation of crime with other factors.

For example, there is information concerning median county income is available from a variety of sources including the US Census Bureau:

http://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?src=CF

##Other parameters

We can study the correlation of crime with other factors.

For example, there is information concerning median county income is available from a variety of sources including the US Census Bureau:

http://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?src=CF

Sadly, the only years available are 2009-2014

##Other parameters

We can study the correlation of crime with other factors.

For example, there is information concerning median county income is available from a variety of sources including the US Census Bureau:

http://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?src=CF

Sadly, the only years available are 2009-2014

We already have the code required to munge and clean the data

##Income information

First get the information into memory:

```{r}
filenames <- list.files("~/Classes/16-17/Fall/EDA2016/Data/IncomeFL/ACS", 
                        pattern="*.csv", full.names=TRUE)

ldf <- lapply(filenames, function(x) 
  read.csv(x, header = TRUE))
ldf <- lapply(ldf, function(x) as.data.frame(x)[,7:9])
Alachua_label <- lapply(ldf, function(x) which((x[,1] == "Alachua County") | (x[,1] =="Alachua") ))
Washington_label <- lapply(ldf, function(x) which((x[,1] == "Washington County") | (x[,1] == "Washington") ))
ldf <- lapply(ldf, NULL_Name)

```

##Income Data Frame
Generate the frame

```{r}
L<- length(as.data.frame(ldf[1])[1,])
FL_Income_0914 <- as.data.frame(ldf[1])[Alachua_label[[1]]:Washington_label[[1]],]

for (i in 2:length(ldf)) {
  FL_Income_0914 <- rbind(FL_Income_0914, as.data.frame(ldf[i])[Alachua_label[[1]]:Washington_label[[1]],])
}

FL_Income_0914 <- as.data.frame(FL_Income_0914)
names(FL_Income_0914) <-c("County", "Median_Income", "SE")
str(FL_Income_0914)
```

##Clean up and write to disk
```{r}
FL_Income_0914[,1] <- as.factor(gsub("\\ County", "", FL_Income_0914[,1]))

start <- rep(2009,67)
FL_Income_0914$Year <- c(start,start+1,start+2,start+3,start+4,start+5)

write.csv(FL_Income_0914, "~/Classes/16-17/Fall/EDA2016/Data/IncomeFL/FL_Income_0914.csv",row.names=FALSE)
```

There is a small problem:  If we want to study the correlation of income and crime, we need *crime rates.*

By now we know how to find such things and construct the required data frame.

##Population data

A good source for population data is the US Census site:

https://www.census.gov/popest/data/counties/asrh/2015/PEPSR6H.html

##Population data

A good source for population data is the US Census site:

https://www.census.gov/popest/data/counties/asrh/2015/PEPSR6H.html


I dowloaded data from 2010-2015.

We load the data and take a look:

```{r}
FL_Pop_2010_15 <- read.csv("~/Classes/16-17/Fall/EDA2016/Data/PopulationFL/PopFL201015.csv", header=TRUE, row.names=NULL)
names(FL_Pop_2010_15)
```

For the time being, "totpop" looks useful

##Total population

It appears we can get what we need by filtering the "sexid" and "hispid" features:

```{r}
FLP <- FL_Pop_2010_15 %>% filter(Sex.id=="totsex" & Hisp.id =="tothisp") %>%select(c(2,9,10))
FLP[,3] <- as.numeric(as.character(FLP[,3]))
labels <- grep("July", FLP[,1])
FLP <- FLP[labels,]
head(FLP,n=10)
```

##Cleaning population

We should get rid of the extra words:

```{r}
FLP[,2] <- gsub("\\ County, Florida", "", FLP[,2])
FLP[,1] <- gsub("July 1, ", "", FLP[,1])
names(FLP) <- c("Year", "County", "Total_pop")
FLP <- FLP %>% arrange(Year)
FLP[,1] <- as.numeric(FLP[,1])
FLP <- FLP[,c(2,3,1)]
FLP[,1] <- as.factor(FLP[,1])
str(FLP)
```

##Correlations

To study correlations between income and crime in Florida, we merge our data frames

```{r}
FCI <- inner_join(FL_Crime_0612, FL_Income_0914)
FCIP <- inner_join(FCI, FLP)
str(FCIP)
```

##Variable(s) of interest
Now we can compute rates.  Let's study assault:

```{r}
FCIPA <- FCIP %>% mutate(Assault_Rate = (Simple.Assault/Total_pop))
```

##Quick images: boxplot
```{r}
ggplot(data=FCIPA, aes(x=as.factor(Year), y=Assault_Rate)) + 
  geom_boxplot()
```

##Quick images: scatterplot
```{r}
ggplot(data=FCIPA, aes(x=Median_Income, y=Assault_Rate)) + 
  geom_point()
```

##Quick images: scatterplot
```{r}
ggplot(data=FCIPA, aes(x=Median_Income, y=Assault_Rate)) + 
  geom_point(aes(color=as.factor(Year)))
```

##Quick images: scatterplot
```{r}
ggplot(data=FCIPA, aes(x=Median_Income, y=Assault_Rate)) + 
  geom_point() +
  geom_smooth(method="lm")
```

##Quick images: scatterplot
```{r}
ggplot(data=FCIPA, aes(x=Median_Income, y=Assault_Rate, group=as.factor(Year))) + 
  geom_point() +
  geom_smooth( method="lm") 
```
