---
title: "EDA Exercise 3"
author: "Aaron Niskin"
date: "August 30, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


University of California Irvine maintains an interesting collection of data sets for those interested in machine
learning

<https://archive.ics.uci.edu/ml/datasets.html>

Navigate to <https://archive.ics.uci.edu/ml/machine-learning-databases/00296/>

and download the zipped diabetes data set.

Prepare an R Markdown document which documents the use of R tools/code to address the following

##1. Read the data into memory.

```{r eval=TRUE}
df <- read.csv("dataset_diabetes/diabetic_data.csv")
```

##2. Document any missing values in the data.

```{r eval=FALSE, echo=FALSE}
### This is data computed using the below function. I stored it as a variable so that I don't hvae to compute that function every time I want to compile the PDF.
invalidRows <- scan(file="invalidRows.txt", sep="\n", what="character")
```

Sorry... I think I'm enjoying this a bit too much. I wanted to try functional programming, so this is me doing that. So, I'm sorry for nerding out super hard on this. But look how cool that is! I can completely reuse that function.

```{r eval=TRUE}
hasInvalidDataNames <- function(dats, nullValues) {
  hasInvalidData <- function(col_name) {
    occurrences <- sapply(dats[,col_name], function(x) {
      is.null(x) || is.na(x) || x %in% nullValues
      })
    length(which(occurrences))
  }
  sapply(names(dats), hasInvalidData)
}
invalidRows_tmp <- hasInvalidDataNames(df, c("", "None", "?", "Unknown/Invalid"))
invalidRows     <- invalidRows_tmp[invalidRows_tmp > 0]
```

```{r}
invalidRows
```


```{r eval=TRUE, echo=FALSE}
### This is so that I don't have to keep on running the above code every time I edit the file.
write.csv(invalidRows, "invalidRows.csv")
```

So, the rows that have missing values (to include any empty strings, question marks, "None"'s, or "Unknown/Invalid"). If one were to decide to not consider one of those values as missing, just recompute with a different second argument vector.

##3. What percentage of patients are admitted from the emergency room? Given a patient is admitted from the emergency room, what is the probability that their discharge status will be “expired”?

So first let's store the subset of those who were admitted from the emergency room, get the count and the total count of all patients admitted into these 150 hospitals.

```{r eval=TRUE}
emergency_admits <- df[df$admission_source_id == 7,]
num_emergency <- dim(emergency_admits)[1]
num_patients <- dim(df)[1]
```

Now, we can divide the number of emergency room admitees by the total number of patients (then multiply by 100) to get the percentage.

```{r eval=TRUE}
answer3a <- num_emergency / num_patients
```

The percentage of people admitted to the hospital from the emergency room is approximately, `r sprintf("%.1f", answer3a * 100)`%.

Now we can further subset the emergencyroom admitees and redo the same.

```{r}
num_exp <- dim(emergency_admits[emergency_admits$discharge_disposition_id == 11,])[1]
answer3b <- num_exp / num_emergency
```

So, the total probability that a person who was admitted from the emergency room will also have an expired discharge status is approximately, `r sprintf("%.1f", answer3b * 100)`%.


##4. What is the most frequent admission status? What is the most frequent discharge status? For the most frequent admission status, what is the most frequent discharge status?

Let's check out our data a bit just to see what we're working with.

```{r}
table(df$admission_type_id)
```

We can see that it's 1 (corresponding to Emergencyroom admittance), but let's do this programmatically:


```{r eval=FALSE, include=FALSE}
getMostFrequentAttr <- function(dats) {
  function(attr) {
    countAttrVal <- function(attrVal) {
      dim(dats[attr==attrVal,])[1]
    }
    sapply(unique(dats[,attr]), countAttrVal)
  }
}

mostFreqAdmStat <-getMostFrequentAttr(df)("admission_type_id")
```

```{r}
admTab <- table(df$admission_type_id)
mostFreqAdm <- names(which(admTab == max(admTab)))
mostFreqAdm
```

And again for disposition id:

```{r}
table(df$discharge_disposition_id)
```

So we should get 1 again (corresponding to "discharged to home").

```{r}
disTabs <- table(df$discharge_disposition_id)
mostFreqDis <- names(which(disTabs == max(disTabs)))
mostFreqDis
```

Just to make sure we're not getting 1 due to some anomaly or misunderstanding of R somewhere, let's try to get a value whose name is not the same as the array index, like 22,1993 by computing:

```{r}
names(which(disTabs == 1993))
```

```{r}
table(df[df$admission_type_id == mostFreqAdm, "discharge_disposition_id"])
```

```{r}
MFATab <- table(df[df$admission_type_id == mostFreqAdm, "discharge_disposition_id"])
names(which(MFATab == max(MFATab)))
```



##5. Characterize the distribution of admission type.

```{r}
hist(df$admission_type_id)
```

It looks roughly like a geometric distribution to me, but I will investigate tools to identify the distribution a bit more so as to get a better estimation.

